---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
auth_success <- tryCatch(
  khisr:::khis_cred_docs(),
  khis_cred_internal_error = function(e) e
)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = TRUE,
  purl = khisr::khis_has_cred(),
  eval = khisr::khis_has_cred()
)

library(khisr)
```

The `khisr` package seamlessly integrates with the Kenya Health Information System 
(KHIS) API from R, empowering researchers and public health professionals to easily 
access and analyze valuable health data. Built on the DHIS2 platform, khisr 
automates data retrieval, saving you time and effort compared to manual methods.

## Authentication

To ensure a secure and seamless experience, `khisr` operates by default in 
authenticated mode, enabling you to interact with KHIS as a recognized user. 
This means that before diving into KHIS data exploration, you'll need to establish 
your credentials with the system.

### Setting Your Credentials:

1. Obtain your KHIS credentials: These typically consist of a username and 
password, which you can acquire through appropriate channels within the KHIS organization.

2. Store your credentials securely: khisr provides a convenient mechanism for 
storing your credentials within your R environment. To learn how to set and 
manage your credentials effectively, please refer to the comprehensive guide at 
[Set Your Credentials](https://khisr.damurka.com/articles/set-your-credentials.html).

```{r, eval = FALSE}
# Set the credentials using username and password
khis_cred(username = 'KHIS username', password = 'KHIS password')

# Set credentials using configuration path
khis_cred(config_path = 'path/to/secret.json')
```

## Metadata

KHIS uses metadata to define the structure and meaning of the data stored in the
system. For more information see [data dimensions](https://khisr.damurka.com/articles/data-dimensions.html)

### Metadata helpers

`khisr` provides a number of high level metadata helpers for obtaining details 
of the different types of metadata. There is a helper for each of the main
metadata categories in DHIS2, in fact auto complete in the users R IDE will often 
make typing these very fast. This is the list of the numerous high level metadata 
helpers in `khisr`.

| khisr function                       | KHIS API Endpoint          |
|:-------------------------------------|:---------------------------|
| `get_categories()`                   | categories                 |
| `get_category_combos()`              | categoryCombos             |
| `get_category_option_combos()`       | categoryOptionCombos       |
| `get_category_option_group_sets()`   | categoryOptionGroupSets    |
| `get_category_option_groups()`       | categoryOptionGroups       |
| `get_category_options()`             | categoryOptions            |
| `get_data_element_group_sets()`      | dataElementGroupSets       |
| `get_data_element_groups()`          | dataElementGroups          |
| `get_data_elements()`                | dataElements               |
| `get_data_sets()`                    | dataSets                   |
| `get_indicator_group_sets()`         | indicatorGroupSets         |
| `get_indicator_groups()`             | indicatorGroups            |
| `get_indicators()`                   | indicators                 |
| `get_option_group_sets()`            | optionGroupSets            |
| `get_option_groups()`                | optionGroups               |
| `get_option_sets()`                  | optionSets                 |
| `get_options()`                      | options                    |
| `get_organisation_unit_groupsets()`  | organisationUnitGroupSets  |
| `get_organisation_unit_groups()`     | organisationUnitGroups     |
| `get_organisation_units()`           | organisationUnits          |
| `get_dimensions()`                   | dimensions                 |
| `get_user_groups()`                  | userGroups                 |
| `get_period_types()`                 | periodTypes                |

### Metadata object filter

To filter the metadata there are several filter operations that can be applied 
to the returned list of metadata. The format of the filter itself is 
straight-forward and follows the pattern property:operator:value, where property 
is the property on the metadata you want to filter on, operator is the comparison 
operator you want to perform and value is the value to check against (not all 
operators require value).

| KHIS Operator   | Infix Operator  | Description                                         |
|:----------------|:----------------|:----------------------------------------------------|
|` eq`            | `%.eq%`         | Equality                                            |
| `!eq`           | `%.~eq%`        | Inequality                                          |
| `ieq`           | `%.ieq%`        | Case insensitive string, match exact                |
| `ne`            | `%.ne%`         | Inequality                                          |
| `like`          | `%.Like%`       | Case sensitive string, match anywhere               |
| `!like`         | `%.~Like%`      | Case sensitive string, not match anywhere           |
| `$like`         | `%.^Like%`      | Case sensitive string, match start                  |
| `!$like`        | `%.~^Like%`     | Case sensitive string, not match start              |
| `like$`         | `%.Like$%`      | Case sensitive string, match end                    |
| `!like$`        | `%.~Like$%`     | Case sensitive string, not match end                |
| `ilike`         | `%.like%`       | Case insensitive string, match anywhere             |
| `!ilike`        | `%.~like%`      | Case insensitive string, not match anywhere         |
| `$ilike`        | `%.^like%`      | Case insensitive string, match start                |
| `!$ilike`       | `%.~^like%`     | Case insensitive string, not match start            |
| `ilike$`        | `%.like$%`      | Case insensitive string, match end                  |
| `!ilike$`       | `%.~like$%`     | Case insensitive string, not match end              |
| `gt`            | `%.gt%`         | Greater than                                        |
| `ge`            | `%.ge%`         | Greater than or equal                               |
| `lt`            | `%.lt%`         | Less than                                           |
| `le`            | `%.le%`         | Less than or equal                                  |
| `token`         | `%.token%`      | Match on multiple tokens in search property         |
| `!token`        | `%.~token%`     | Not match on multiple tokens in search property     |
| `in`            | `%.in%`         | Find objects matching 1 or more values              |
| `!in`           | `%.~in%`        | Find objects not matching 1 or more values          |

### Working with metadata filters

Basic usage of the metadata filter

```{r}
# Retrieve organisation units by county (level 2)
county <- get_organisation_units(level %.eq% '2')
county

# Retrieve county by name (Mombasa)
county <- get_organisation_units(level %.eq% '2',
                                 name %.like% 'mombasa')
county

data_element_id <- c('cXe64Yk0QMY', 'XEX93uLsAm2')

# Retrieve data elements by ID using operator in
data_elements <- get_data_elements(id %.in% data_element_id)
data_elements

# Retrieve data elements by filtering using dataElementGroups
data_elements <- get_data_elements(dataElementGroups.name %.like% 'moh 705')
data_elements
```

## Data analytics

The analytics resource in DHIS2 empowers you to access and analyze aggregated data
across various dimensions. To effectively leverage this resource, let's explore the
key functions and parameters involved:

### Key Functions

- `get_analytics()`: Retrieves aggregated data based on specified dimensions and
  filters.
- `analytics_dimension()`: Constructs dimensions for queries, ensuring accurate
  data retrieval.
- `%.d%` (infix operator): Convenient shorthand for creating dimension filters.
- `%.f%` (infix operator): Convenient shorthand for creating filter dimensions.

### Dimension (dx)

The `dimension` query parameter defines which dimensions should be included in the
analytics query. Any number of dimensions can be specified. The dimension parameter
should be repeated for each dimension to include in the query response. The query
response can potentially contain aggregated values for all combinations of the
specified dimension items. The fixed dimensions are the **data element** _(dx)_
**period (time)** _(pe)_ and **organisation unit**  _(ou)_ dimension. You can 
dynamically add dimensions through categories, data element group sets and 
organisation unit group sets.

| Dimension ID | Dimensions                           |
|:-------------|:-----------------------------------------------------------------|
| dx           | Data elements, indicators, data set reporting rate metrics,      |
|              | data element operands, program indicators, program data elements,|
|              | program attributes, validation rules                             |
|              |                                                                  |
| pe           | ISO periods and relative periods (see "date and period format")  |
|              |                                                                  |
| ou           | Organisation unit hierarchy                                      |
|              | Organisation unit identifiers, keywords USER_ORGUNIT,            |
|              | USER_ORGUNIT_CHILDREN, USER_ORGUNIT_GRANDCHILDREN, LEVEL-<level>,|
|              | and OU_GROUP-<group-id>                                          |
|              |                                                                  |
| co           | Category option combo identifiers (use `all` to get all items)   |
|              |                                                                  |
| ao           |Category option combo identifiers (use `all` to get all items)    |


### Filter (filter)

The `filter` parameter defines which dimensions should be used as filters for the
data retrieved in the analytics query. Any number of filters can be specified. The
filter parameter should be repeated for each filter to use in the query. A filter
differs from a dimension in that the filter dimensions will not be part of the
query response content, and that the aggregated values in the response will be
collapsed on the filter dimensions. In other words, the data in the response will
be aggregated on the filter dimensions, but the filters will not be included as
dimensions in the actual response.

### Constructing Queries

- **Specify Dimensions:** Use the dimensions parameter to define the dimensions 
  you want to include in the query response. Leverage the infix operators %.d% for 
  concise and readable code.

```{r}
# To include a list dimensions for data elements id, dataset ids
dx %.d% c('dimension-id-1', 'dimension-id-2')

pe %.d% 'LAST_YEAR'

ou %.d% 'USER_ORGUNIT'

# showing in the analytics
get_analytics(
    dx %.d% c('siOyOiOJpI8', 'Lt0FqtnHraW', 'OoakJhWiyZp'),
    pe %.d% 'LAST_YEAR',
    ou %.d% c('qKzosKQPl6G')
)

# Using the startDate and endDate with organisation unit keyword 'USER_ORGUNIT'
get_analytics(
    dx %.d% c('siOyOiOJpI8', 'Lt0FqtnHraW', 'OoakJhWiyZp'),
    ou %.d% 'USER_ORGUNIT',
    pe %.d% 'all',
    startDate = '2023-07-01',
    endDate = '2023-12-31'
)
```

- **Apply Filters:** Use the filters parameter to specify dimensions for filtering
  data without including them in the response.
  
```{r}
# Filter by period
pe %.f% 'LAST_YEAR'

# Filter by organisation unit
ou %.f% 'USER_ORGUNIT'

# showing in the analytics. filter by organisation unit with id 'qKzosKQPl6G'
# and period 'LAST_YEAR'
get_analytics(
    dx %.d% c('siOyOiOJpI8', 'Lt0FqtnHraW', 'OoakJhWiyZp'),
    pe %.f% 'LAST_YEAR',
    ou %.f% 'qKzosKQPl6G'
)
```