---
title: "Tracker Data"
---

```{r, include = FALSE}
can_decrypt <- httr2::secret_has_key('KHIS_KEY')

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = TRUE,
  purl = can_decrypt,
  eval = can_decrypt
)
options(tibble.print_min = 5L, tibble.print_max = 5L)
```

```{r, eval = !can_decrypt, echo = FALSE, comment = NA}
message("No token available. Code chunks will not be evaluated.")
```

```{r, eval = can_decrypt, include = FALSE}
khisr:::khis_cred_docs()
```

```{r, include = FALSE}
library(khisr)
```

# Overview

Alongside the [aggregate data](https://khisr.damurka.com/articles/data-dimensions.html)
covered by `get_analytics()`, DHIS2's **Tracker** stores case-based, person-level
data — the kind of record you'd use for individual patients or clients rather than
facility-level totals. `khisr` provides three functions for reading it:

| khisr function            | DHIS2 API Endpoint         | Retrieves                          |
|:---------------------------|:----------------------------|:------------------------------------|
| `get_tracked_entities()`   | `tracker/trackedEntities`   | The people/subjects being tracked  |
| `get_enrollments()`        | `tracker/enrollments`       | Their enrollment in a program      |
| `get_events()`             | `tracker/events`            | What happened at each program stage visit |

These map onto four core Tracker concepts:

| Concept            | Description                                                                                     |
|:--------------------|:--------------------------------------------------------------------------------------------------|
| **Tracked Entity**  | The subject being followed over time (e.g. a patient or client). Identified by a `trackedEntityType` such as "Person". |
| **Program**         | The case-based program a tracked entity is enrolled in (e.g. a child health programme).          |
| **Enrollment**      | A tracked entity's participation in a specific program, with its own enrollment and incident dates. |
| **Event**           | A single visit/data-collection point within an enrollment, tied to one `programStage` (e.g. a birth visit) and carrying `dataValues`. |

All three functions are `r lifecycle::badge("experimental")` and share the same
scoping, pagination, and error-handling behaviour.

# Scoping a query

Every Tracker query must be scoped — DHIS2 rejects an unscoped request. Scope by
`program`, and/or by `org_units` together with `org_unit_mode`, which controls how
`org_units` is interpreted:

| `org_unit_mode` | Meaning                                                          |
|:------------------|:--------------------------------------------------------------------|
| `SELECTED`         | Only the given org units (DHIS2's default when `org_units` is set)  |
| `CHILDREN`         | The given org units and their immediate children                    |
| `DESCENDANTS`      | The given org units and all units below them in the hierarchy       |
| `ACCESSIBLE`       | Everything the authenticated user has access to (default when `org_units` is omitted) |
| `CAPTURE`          | Everything the authenticated user can capture data into             |
| `ALL`              | Every org unit in the instance (typically requires elevated authority) |

```{r}
# All tracked entities enrolled in a program, at an org unit and everything below it
get_tracked_entities(
    program = 'IpHINAT79UW',
    org_units = 'DiszpKrYNg8',
    org_unit_mode = 'DESCENDANTS'
)
```

# Retrieving tracked entities

`get_tracked_entities()` returns one row per tracked entity. Repeating structures
such as `attributes` and `enrollments` come back as list-columns — use
[tidyr::unnest_wider()] or [tidyr::unnest_longer()] to flatten the parts you need.

```{r}
entities <- get_tracked_entities(
    program = 'IpHINAT79UW',
    org_units = 'DiszpKrYNg8',
    org_unit_mode = 'DESCENDANTS',
    fields = c('trackedEntity', 'orgUnit', 'attributes')
)
entities
```

## Filtering by attribute value

`get_tracked_entities()` also accepts one or more [tracked_entity_filter()]
attribute filters. This is a separate, smaller operator set from
[metadata_filter()] — it applies only to the `trackedEntities` endpoint, and the
`in` operator joins values with semicolons rather than a comma-separated bracket
list:

| Operator | Meaning                     |
|:----------|:------------------------------|
| `eq`      | Equality                     |
| `ne`      | Inequality                   |
| `ge`      | Greater than or equal        |
| `gt`      | Greater than                 |
| `le`      | Less than or equal           |
| `lt`      | Less than                    |
| `like`    | Match anywhere               |
| `sw`      | Starts with                  |
| `ew`      | Ends with                    |
| `in`      | Match one or more values     |
| `null`    | Attribute has no value       |
| `!null`   | Attribute has a value        |

```{r}
# Tracked entities whose attribute w75KJ2mc4zz starts with "Ann"
get_tracked_entities(
    program = 'IpHINAT79UW',
    org_units = 'DiszpKrYNg8',
    org_unit_mode = 'DESCENDANTS',
    tracked_entity_filter('w75KJ2mc4zz', 'sw', 'Ann')
)
```

# Retrieving events

`get_events()` returns one row per event, with the data collected at that visit
in a `dataValues` list-column.

```{r}
get_events(
    program = 'IpHINAT79UW',
    org_units = 'DiszpKrYNg8',
    org_unit_mode = 'DESCENDANTS'
)
```

DHIS2's Tracker API does not currently document a way to filter events by data
element value the way `trackedEntities` supports attribute filtering, so
`get_events()` has no `filter` argument. If your instance supports this via an
undocumented parameter, pass it through `...`; otherwise retrieve the broader
event set and filter client-side after unnesting `dataValues`.

# Retrieving enrollments

`get_enrollments()` returns one row per enrollment, useful for questions about
who is enrolled and when, independent of any particular visit.

```{r}
get_enrollments(
    program = 'IpHINAT79UW',
    org_units = 'DiszpKrYNg8',
    org_unit_mode = 'DESCENDANTS'
)
```

# Date filters and pagination

All three functions accept `updated_after`/`updated_before`, and
`get_tracked_entities()`/`get_enrollments()` additionally accept
`enrolled_after`/`enrolled_before` and `occurred_after`/`occurred_before` to
bound results by enrollment or incident date. `get_events()` accepts
`occurred_after`/`occurred_before` for the event's own occurred date.

Results are paginated automatically — `page_size` (default 500) controls how many
records are requested per page, not the total returned; all matching pages are
fetched and combined into a single tibble.

# Further resources

For the full set of query parameters your DHIS2 instance supports, see the
[DHIS2 Tracker API documentation](https://docs.dhis2.org/en/develop/using-the-api/dhis-core-version-master/tracker.html).
